---
title: "enrichmentHeatmap - Fig4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
library(tidyverse)
library(readxl)
```


Defines functions that read each enrichment file in a specific folder, concats all files in one big table and adds one extra column that specifies the tissue name. 
```{r }
read_concat_files = function(myFolder) {
  filenames <- list.files(myFolder, pattern="*.xlsx", full.names=TRUE)
  tibs <- lapply(filenames, read_xls_files)
  for(i in 1:(length(tibs)-1)) { 
    tibs[[i+1]] = dplyr::bind_rows(tibs[[i]], tibs[[i+1]])
    }
  return(tibs[[length(tibs)]])
}

#provides functionality to read multiple tissue enrichment files
read_xls_files = function(file) {
  f <- str_split(file, "/", simplify = TRUE)
  n <- length(f)
  df <- read_excel(file, col_types= c("text","text","numeric","numeric","numeric","numeric","numeric","numeric","numeric")) %>% mutate(tissue = str_split(f[n], "_TiP", simplify = TRUE)[1])
  return(df)
}

```


Get the matrix with all TiP neighbors per tissue (exported from the python notebook)
TiPneighbors_to_PEX_alltissues_custom.csv

```{r}
dir <- "../Manuskript/"

my_tipNeighbors_file <- paste(dir, "Julia_Analysen/results/parameter/TiPneighbors_to_PEX_alltissues_customsymbol.csv", sep="")
tipNeighbors <- read_csv(my_tipNeighbors_file)
#turn data into tidy format and remove all unintersting tissues eg those with NA
tipNeighbors <- rename(tipNeighbors, "TiP" = ...1) %>% 
  pivot_longer(cols = 2:56, names_to = "tissue", values_to = "pex_interactor") %>% 
  filter(!is.na(pex_interactor)) %>% 
  mutate(pex_interactor = str_replace_all(pex_interactor,"\\['|'\\]|'","")) 

#show me a list of TiP and count how often they are TiP and neighbor of pex proteins
tipNeighbors %>% 
  count(TiP) %>% 
  arrange(desc(n)) 
```
Plot n of TiP neighbors per tissue
```{r}
bp <- tipNeighbors %>% 
  count(tissue) %>% 
  ggplot(aes(x = reorder(tissue, desc(n)), y = n)) + 
  geom_bar(stat = "identity") + 
  theme_light() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(colour = "grey", linetype ="dotted")) +
  xlab("tissue") + 
  ylab("n of interacting TiP")

my_file <- paste(dir, "Figures/TiPperTissue.pdf", sep="")
ggsave(my_file, plot = bp)

bp

```

```{r}
#show n of TiPs 
tipNeighbors %>% rowwise() %>% mutate(n_interactors = length(unlist(str_split(pex_interactor, ",")))) %>% ungroup %>% arrange(desc(n_interactors))
```
OK. not surprisingly there is UBC. Lets plot a heatmap to get a better overview
```{r}
#show n of TiPs and > 1 pex interactors
tmp_neighbors <- tipNeighbors %>% 
  rowwise() %>% 
  mutate(n_interactors = length(unlist(str_split(pex_interactor, ",")))) %>% 
  ungroup %>%
  filter(n_interactors > 1)

tmp <- tmp_neighbors %>%
  select(TiP, tissue, n_interactors) %>% 
  pivot_wider(names_from = tissue, values_from = n_interactors) %>%
  mutate_all(~replace(.,is.na(.), 0))

matrix <- as.matrix(column_to_rownames(tmp, var = "TiP")) 
#cluster TiP
ord_tip <- hclust( dist(matrix, method = "euclidean"))$order
#ord
#cluster Tissues
ord_tissues <- hclust( dist(t(matrix), method = "euclidean"))$order

hp <-  tmp_neighbors %>% 
  mutate(TiP = factor(TiP, levels = rownames(matrix)[ord_tip])) %>%
  mutate(tissue = factor(tissue, levels = colnames(matrix)[ord_tissues])) %>%
  ggplot(aes(x = tissue, y = TiP)) + 
  geom_tile(aes(fill = n_interactors)) +
  theme_light() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),panel.grid.major.x = element_line(colour = "grey", linetype ="dotted"), panel.grid.major.y = element_line(colour = "grey", linetype ="dotted")) + xlab("") + ylab("")

my_file <- paste(dir, "Figures/heatmapTiP_numberInteractors_perTissue.pdf", sep="")
# printing (cm)
A4width = 21
A4height = 29.7
ggsave(my_file, plot = hp, width = 0.35*A4width, height = 0.6*A4height)

rm(tmp_neighbors, tmp, matrix)
hp
  
```




Plot enriched GO Terms of neighboring TiPs

Run functions to prepare data for plotting
```{r warning= FALSE}
#set the desired GO-ontology here eg MF, BP or CC
ontology <- "BP"

my_folder <- paste(dir, "Julia_Analysen/results/parameter/enrichment/all_enrich_julia/", ontology, "/", sep = "")
my_heatmap_file <- paste(dir, "Figures/Figure4/expressionEnrichmentHeatmap_",ontology,".pdf", sep = "")

dat_enr <- read_concat_files(my_folder)

```

Run a clustering to reorder tissues
```{r}
#filter for sign results and show only those terms that have > 1 genes 
dat_enr %>% 
  filter(`P-value-corrected`< 0.05) %>% 
  filter(Abundance_group > 1) %>% 
  select(tissue, Association_term, `P-value-corrected`, OR) -> clustdata

clustdata %>% 
  select(tissue, Association_term, `P-value-corrected`) -> temp
#prepare tibble for conversion into matrix
temp %>% 
  pivot_wider(names_from=c(Association_term), values_from= c('P-value-corrected')) -> clustdata_wide
rm(temp)
clustdata_wide %>% 
  mutate_all(~replace(.,is.na(.), 0)) -> temp
#convert to matrix
as.matrix(column_to_rownames(temp, var = "tissue")) -> matrix
rm(temp)

#cluster tissues
ord <- hclust( dist(matrix, method = "euclidean"))$order
#ord
#cluster go terms
ord_go <- hclust( dist(t(matrix), method = "euclidean"))$order
ord_go
```



Now we plot the heatmap and save it to file "my_heatmap_file" as pdf
```{r}
# defines colours used in the heatmap -> color brewer but reversed order, since we want lower p-values to be "blueer"
cols = rev(RColorBrewer::brewer.pal(9, "Blues"))
#remove white from the cols and add a light-grey
cols = cols[-9]
cols[9] = "#f0f0f0" 

p <- clustdata %>% 
  #mutate(Association_term = factor(Association_term, levels = colnames(matrix)[ord_go])) %>% 
  mutate(tissue = factor(tissue, levels = rownames(matrix)[ord])) %>%
  ggplot(aes(x=tissue, y=Association_term)) + 
  geom_tile(fill= "white") + 
  theme_light() + 
  theme(panel.spacing = unit(0.25, "lines"),axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),panel.grid.major.y = element_line(colour = "grey", linetype ="dotted")) + 
  geom_point(aes(size = OR, fill = `P-value-corrected`, colour = `P-value-corrected`), alpha = 0.7) +
  scale_fill_gradientn(colors = (cols), limits = c(0,0.05)) + 
  scale_size(range = c(1,6)) +
  coord_flip() +
  labs(x = "", y = "") + 
  guides(fill = "none")

# printing (cm)
A4width = 21
A4height = 29.7
  
ggsave(my_heatmap_file, plot = p, width = 0.4*A4width, height = 0.175*A4height, scale = 1.5)

p
```



