---
title: "Figures 1&2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Imports
```{r}
library(tidyverse)
library(readxl)
library(ggpubr)
library(rstatix)
library(igraph)
library(readxl)
```


Variables and data
```{r}
#colors from iwanthue https://medialab.github.io/iwanthue/?utm_source=Nature+Briefing&utm_campaign=e0c3e024d4-briefing-dy-20211006&utm_medium=email&utm_term=0_c9dfd39373-e0c3e024d4-44542065
colors <- read_tsv("data/colors_fancy.txt", col_names = F, show_col_types = FALSE) %>% pull()
mygrey <- "#F5F5F5"

#import
my_file_pos <- "data/2_Bret_screen_final.xlsx"
#literature interactions
my_file_interactions <- "data/Peroxisomal_interactions_across_dbs_final.xlsx" 
#import of Yifrach Dataset
#Gene / FinalScore / DetectionFrequency
my_yifrach_genes <- "data/yifrach.xlsx"
#ePI interactome
ePI_file <- "data/ePI_edgelist.csv"
#random sampling lcc
my_random_lcc_file <- "data/1000_degree_matched_random_samples_LCC_disease_genes_in_interactome.xlsx"
#data on degree of ePI members
my_intExt_degree_file <- ("../Manuskript/Julia_Analysen/173_ePI/results/parameter/characteristics_ePI_degrees.csv")
#contains comparison of lit vs bret interactions
my_file_mapping <- "data/mapper.xlsx" #mapping from entrez to symbol
my_file_function <- "data/9_genesFunction.txt"  #functions of peroxisomal proteins

#output
my_heatmap_file = "output/figure2/2B_heatmap_positives.pdf" #output for heatmap file
```


Genes from the results
```{r}
positives_original <- read_xlsx(my_file_pos, col_names=TRUE)
g_isoforms <- sort(union(unique(positives_original$ProteinA), unique(positives_original$ProteinB)))
g_isoforms
```

Prepare everything for heatmap plotting
```{r}
#read all positives
positives <- read_xlsx(my_file_interactions) %>% 
  replace(is.na(.), 0)

#load mapper and prepare 
mapper <- read_excel(my_file_mapping)
mapper_red <- mapper %>% 
  dplyr::select('Symbol[table]', 'EntrezID[mapper]') %>% 
  group_by(`Symbol[table]`,`EntrezID[mapper]`) %>% 
  distinct() %>% 
  ungroup()

#map data
#code data
#found in ucrinteractome = 1
#found in iBRET = 2
#found in both = 3
#found nothing = 0
positives_mapped <- positives %>% 
  left_join(mapper_red, by = c("Interactor_A" = "EntrezID[mapper]")) %>% 
  left_join(mapper_red, by = c("Interactor_B" = "EntrezID[mapper]")) %>% 
  arrange(Interactor_A) %>% 
  dplyr::rename(proteinA = 'Symbol[table].x', proteinB = 'Symbol[table].y') %>% 
  mutate(class = if_else(InteractomeUCR == 1,if_else(iBret == 1,3,1),if_else(iBret == 1,2,0)))

positives_mapped %>% group_by(class) %>% count()
```


```{r}
#do we have all proteins? n = 92 (without isoforms)
pg_withoutIso <- sort(union(unique(positives_mapped$proteinB), unique(positives_mapped$proteinA)))

pg_withoutIso
```

################################
JACCARD Similarity for isoforms
Are their interaction partners more similar than other proteins?

Lets get the isoforms that are included in the screen
```{r}
setdiff(g_isoforms, pg_withoutIso) -> i
setdiff(i, "ABCD3b") -> isos #we do not have a partner for ABCD3b -> exclude it here
isos
```

```{r}
p_inter_with_iso <- positives_original %>% 
  filter(!ProteinA %in% remove_genes & !ProteinB %in% remove_genes) %>% 
  select(ProteinA, ProteinB)

p_graph_with_iso <- graph_from_edgelist(as.matrix(p_inter_with_iso), directed = FALSE)
#create graph without the isoforms in order to avoid redundancy
p_graph_without_iso <- p_graph_with_iso - isos

#jaccard similarity for the 6 isoforms investigated
#mean((colSums(jacc_matr)-1)/(nrow(jacc_matr)-1))

#isoform_pairs
accaa1 <- c("ACAA1a","ACAA1b")
gstk1 <- c("GSTK1a","GSTK1b")
pex10 <- c("PEX10v1","PEX10v2")
pex5 <- c("PEX5b","PEX5d")
scp2 <- c("SCP2v1","SCP2v2")
sod2 <- c("SOD2v2","SOD2vX" )

isoform_pairs <- list(accaa1, gstk1, pex10, pex5, scp2, sod2)

#calculate jaccard similarity for each pair of isoforms
jacc_isoforms <- sapply(isoform_pairs, function(x) similarity(p_graph_with_iso, x, method="jaccard"))

#jaccard similarity for all other pairs without isoforms
jacc_allother <- similarity(p_graph_without_iso, method="jaccard")
jacc_tbl <- tibble(jaccard = (colSums(jacc_allother)-1)/(nrow(jacc_allother)-1), cat = "other")

#bind pairwise jaccsimilarity of isoforms and others together
jacc_all <- rbind(jacc_tbl, tibble(jaccard = jacc_isoforms[2,], cat = "isoforms"))

jaccard_plot <- jacc_all %>% 
  ggplot(aes(x = cat, y=jaccard)) + 
  geom_boxplot(aes(fill = cat)) + 
  theme_classic() +
  labs(x = "", y = "") + 
  scale_fill_manual(values = c("#94d9c5", "#D3D3D3")) + stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 0.3, label = "p.format") + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none")

ggsave("output/figure1/jaccard_isoform_boxplot.pdf", plot = jaccard_plot, width = 3, height = 5, units = "cm", scale = 1.6)
jaccard_plot
```


```{r}
jacc_all %>% 
  group_by(cat) %>% 
  summarise(mean = mean(jaccard), median = median(jaccard), sd = sd(jaccard))
```

Is the higher Jaccard Index for isoforms driven by PPI to the import receptors PEX5, PEX7, PEX19?

```{r}
#lets remove PPI to PEX5, PEX7 and PEX19 and repeat analysis of Jaccard Index
p_graph_with_iso_withoutImport <- p_graph_with_iso - c("PEX5b", "PEX5d", "PEX7", "PEX19") 
p_graph_without_iso_withoutImport <- p_graph_without_iso - c("PEX7", "PEX19") 

jacc_isoforms_withoutImport <- sapply(list(accaa1, gstk1, pex10, scp2, sod2), function(x) similarity(p_graph_with_iso_withoutImport, x, method="jaccard"))

jacc_allother_withoutImport <- similarity(p_graph_without_iso_withoutImport, method="jaccard")

jacc_all_withoutImport <- rbind(jacc_tbl_withoutImport, tibble(jaccard = jacc_isoforms_withoutImport[2,], cat = "isoforms"))

jaccard_plot <- jacc_all_withoutImport %>% 
  ggplot(aes(x = cat, y=jaccard)) + 
  geom_boxplot(aes(fill = cat)) + 
  theme_classic() +
  labs(x = "", y = "") + 
  scale_fill_manual(values = c("#94d9c5", "#D3D3D3")) + stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 0.3, label = "p.format") + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none")

#ggsave("../Manuskript/Figures/Figure1/jaccard_isoform_boxplot.pdf", plot = jaccard_plot, width = 3, height = 5, units = "cm", scale = 1.6)
print(jaccard_plot)

```



```{r}
jacc_all_withoutImport %>% 
  group_by(cat) %>% 
  summarise(mean = mean(jaccard), median = median(jaccard), sd = sd(jaccard))

```



PPI Matrix for isoforms
```{r}
isoform_pairs
getIsoformPPIforMatrix <- function(isoform, g) {
  return( tibble(isoform = isoform, interactors = neighbors(g, v= isoform)$name) %>% distinct())
}

isoform_ppi <- lapply(isos, function(x) getIsoformPPIforMatrix(x, p_graph_with_iso)) %>% bind_rows %>% mutate(class = "ppi")

p<- isoform_ppi %>% ggplot() + 
  geom_tile(aes(x = interactors, y = isoform, fill = class), color="white") + 
  scale_fill_manual(values=colors[4], name="", na.value=mygrey) +
  theme_light() + 
  theme(panel.spacing = unit(0.5, "lines"),axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),panel.grid.major.y = element_line(colour = "grey", linetype ="dotted")) + 
  labs(x="", y="") + 
  guides(fill = "none")

ggsave("output/figure1/isoform_ppi_matrix.pdf", plot = p, width = 13, height=4, scale = 1.6, units = "cm")

print(p)
```



Comparison with yifrach dataset
```{r}
yifrach_genes <- read_excel(my_yifrach_genes)

yifrach_genes %>% 
  mutate(study = if_else(Gene %in% pg_withoutIso,"y","n")) %>% 
  group_by(study, DetectionFrequency) %>% 
  count()
```

Heatmap of ePI (iBRET positives plus literature interactions of all pex associated genes according to Yifrach et al)

(1) Read ePI interactions
```{r}
ePI_edge <- read_csv(ePI_file, show_col_types = FALSE) %>% 
  dplyr::select(Symbol_A, Symbol_B) %>% 
  dplyr::rename(proteinA = "Symbol_A", proteinB = "Symbol_B")

ePI_graph <- graph_from_edgelist(as.matrix(ePI_edge), directed = FALSE)
```

(2)
some operations that prepare the data
specifically, this sets a lower triangle for heatmap plotting
```{r}
#create igraph with bret interactions only
pex_bret <- graph_from_edgelist(as.matrix(positives_mapped %>% filter(class == 2) %>% dplyr::select(proteinA, proteinB)), directed = FALSE)
#create igraph with literature interactions only
pex_lit <- graph_from_edgelist(as.matrix(positives_mapped %>% filter(class == 1) %>% dplyr::select(proteinA, proteinB)), directed = FALSE)
#create igraph that contains only ppi detected by ibret and already known
pex_both <- graph_from_edgelist(as.matrix(positives_mapped %>% filter(class == 3) %>% dplyr::select(proteinA, proteinB)), directed = FALSE)

#add class eg whether found with ibret or literature evidence or both
inter_both <- as_data_frame(igraph::intersection(ePI_graph, pex_both, keep.all.vertices = FALSE)) %>% mutate(class = 3)
inter_bret <- as_data_frame(igraph::intersection(ePI_graph, pex_bret, keep.all.vertices = FALSE)) %>% mutate(class = 2)
inter_lit <- as_data_frame(igraph::intersection(ePI_graph, pex_lit, keep.all.vertices = FALSE)) %>% mutate(class = 1)
inter_ePI <- as_data_frame(difference(ePI_graph, union(pex_bret, pex_both, pex_lit))) %>% unique() %>% mutate(class = 1)
ePI_edge_cl <- rbind(inter_both, inter_bret, inter_lit, inter_ePI) %>% dplyr::rename(proteinA = "from", proteinB = "to")
rm(inter_both,inter_bret, inter_lit, inter_ePI)

ePI_edge_cl -> posA
ePI_edge_cl %>% dplyr::select(proteinB, proteinA, class) %>% rename(proteinA = "proteinB", proteinB = "proteinA") -> posB

#bind everything together and create temp. dataframe. 
df = rbind(posA, posB)

#create a matrix out of the data frame
df_rearranged <- df %>% group_by(proteinA, proteinB) %>% unique %>% ungroup %>% arrange(proteinA) %>% pivot_wider(names_from = proteinA, values_from = class) %>% arrange(proteinB)

df_matr <- as.matrix(dplyr::select(df_rearranged,-proteinB)) 
rownames(df_matr) <- df_rearranged$proteinB

#convert back to data frame but only the lower triangle
ut <- lower.tri(df_matr, diag = TRUE)
df_lower <- data.frame(proteinB = rownames(df_matr)[row(df_matr)[ut]], proteinA = colnames(df_matr)[col(df_matr)[ut]], class = df_matr[ut])

#clean
rm(df)
```

Heatmap plotting
```{r}
#we need the degrees to plot only the top n hub proteins
ePI_degree <- as_tibble(rownames_to_column(as.data.frame(degree(ePI_graph, loops = FALSE)), var="symbol")) %>% dplyr::rename("degree" = `degree(ePI_graph, loops = FALSE)`) 
prot_to_plot <- ePI_degree %>% arrange(desc(degree)) %>% head(n=20) %>% pull(symbol)

p <- df_lower %>% 
  #filter(!is.na(class)) %>%
  mutate(class = case_match(class, 1 ~ "Literature", 2 ~ "iBRET", 3 ~ "Both", NA ~ "not interacting")) %>% 
  mutate(proteinBrev = factor(proteinB, levels = rev(unique(df_lower$proteinB)))) %>%
  ggplot() +
  geom_tile(aes(x = proteinA, y = proteinBrev, fill= class), color="white") +
  theme_light(base_size = 8) +
  scale_fill_manual(values=c("Both" = colors[5], "iBRET" = colors[4], "Literature" = colors[3], "not interacting" = mygrey), name="") +
  #scale_x_discrete(breaks = prot_to_plot) + 
  #scale_y_discrete(breaks = prot_to_plot) +
    theme(
        panel.border = element_rect(linetype = "blank", fill = NA),
        panel.grid = element_line(colour = "white", linetype = "dotted"),
        panel.spacing = unit(0.3, "lines"),
        axis.text.x = element_text(
            angle = 90,
            hjust = 1,
            vjust = 0.5
        ),
    ) +
    labs(x = "", y = "") +
    guides(fill = "none")

ggsave(my_heatmap_file, plot = p, width = 14.5, height=14.5, scale = 1.4, units = "cm")
p
```


Fig 1D
Barplot Top 10 Hubs
```{r}
#exclude literature interactions
pex_graph <- graph_from_edgelist(as.matrix(positives_mapped %>% filter(class != 1) %>% 
                                             dplyr::select(proteinA, proteinB)), directed = FALSE)
pex_degree <- as_tibble(rownames_to_column(as.data.frame(degree(pex_graph, loops = FALSE)), var="symbol")) %>% dplyr::rename("degree" = `degree(pex_graph, loops = FALSE)`) 

p_bar <- pex_degree %>%
  mutate(ranking = dense_rank(desc(degree))) %>%
  filter(ranking < 8) %>%
  ggplot(aes(x = reorder(symbol, degree), y = degree)) + 
  geom_bar(stat = "identity", fill="#D3D3D3", color="black", width=0.7, linewidth=0.3) + 
  theme_minimal() + 
  theme(panel.border = element_rect(linetype = "blank", fill = NA),
        panel.grid = element_line(colour = "white"), 
        panel.grid.major.x = element_line(colour = "grey")) +
  scale_y_continuous(limits = c(0,60)) +
  labs(x = "", y = "") + 
  coord_flip()
p_bar

ggsave("output/figure1/top_hubs.pdf", plot = p_bar, width = 3.5, height = 3.5, units = "cm", scale = 1.5)
```

Fig 1C degree distribution
```{r}
# List of degrees
G.degrees <- degree(pex_graph)

# Let's count the frequencies of each degree
G.degree.histogram <- as.data.frame(table(G.degrees))

# Need to convert the first column to numbers, otherwise
# log-log will not work
G.degree.histogram[,1] <- as.numeric(G.degree.histogram[,1])

# Now, plot it!
p <- ggplot(G.degree.histogram, aes(x = log(G.degrees), y = log(Freq))) +
  geom_point(shape = 21, colour = "black", fill = "grey", stroke = 1) +
  scale_x_continuous("Number of interaction partners") +
  scale_y_continuous("Number of proteins", breaks = c(0,1,2,3)) +
  theme_classic()
ggsave("output/figure1/degree_distribution.pdf", plot = p, width = 4, height = 3, units = "cm", scale = 2)
p

```

Check for power law distribution
```{r}
library(poweRlaw)

#deg <- G.degrees[G.degrees > 0] 
deg <- G.degrees_ePI[G.degrees_ePI > 0] 

pl_model <- displ$new(deg)  # discrete power law

#xmin
est_pl <- estimate_xmin(pl_model)
pl_model$setXmin(est_pl)

#alpha
pl_model$setPars(est_pl$pars)
pl_model$pars

#goodness of fit
bs_pl <- bootstrap_p(pl_model, no_of_sims = 1000, threads = 2)
bs_pl$p
#p ≥ 0,1 → Power law plausible
#p < 0,05 → Power law unlikely

```

Compare to log-normal distribution
```{r}
ln_model <- dislnorm$new(deg)
#ximn
ln_model$setXmin(pl_model$getXmin())
ln_model$setPars(estimate_pars(ln_model))

#likelihood test
comp <- compare_distributions(pl_model, ln_model)
#comp

#results
comp$test_statistic
comp$p_two_sided

# Save as PNG using png() and dev.off()
#png(paste(path_to, "PI_distr_comparison.png", sep=""))
png("output/figure2/ePI_distr_comparison.png")


#plot both models (eg powerlaw vs log-normal)
plot(pl_model, main = "Degree distribution")
lines(pl_model, col = "red")
lines(ln_model, col = "blue")

legend("bottomleft",
       legend = c("Power law", "Log-normal"),
       col = c("red", "blue"),
       lty = 1)



# Close the PNG device
dev.off()


```


Load edgelist of extendedPI and plot degree distribution
```{r}
# List of degrees
G.degrees_ePI <- degree(ePI_graph)

# Let's count the frequencies of each degree
G.degree_ePI.histogram <- as.data.frame(table(G.degrees_ePI))

# Need to convert the first column to numbers, otherwise
# the log-log thing will not work (that's fair...)
G.degree_ePI.histogram[,1] <- as.numeric(G.degree_ePI.histogram[,1])

# Now, plot it!
p <- ggplot(G.degree_ePI.histogram, aes(x = log(G.degrees_ePI), y = log(Freq))) +
  geom_point(shape = 21, colour = "black", fill = "grey", stroke = 1) +
  scale_x_continuous("Number of interaction partners") +
  scale_y_continuous("Number of proteins", breaks = c(0,1,2,3)) +
  theme_classic()
ggsave("output/figure2/degree_distributio_ePI.pdf", plot = p, width = 4, height = 3, units = "cm", scale = 2)
p
```


Figure 2F: Create Barplot for random sampling
```{r}
p<- read_xlsx(my_random_lcc_file) %>% 
  dplyr::rename(lcc = "LCC size of random disease genes") %>% 
  ggplot(aes(x=lcc)) + 
  geom_histogram(binwidth=1, color="white", fill="#D3D3D3") + 
  geom_vline(aes(xintercept = mean(lcc)), color="black", linetype="dashed") +
  theme_minimal() +
  scale_x_continuous(limits = c(0, 60)) +
  xlab("lcc size") +
  ylab("Number of permutations")

ggsave("output/figure2/lcc_disease_module_histogram.pdf", plot = p, width = 4.5, height = 3, units = "cm", scale = 2) 

p

```



Read functions 
```{r}
#read genes with their respective functions and sort
pex_functions <- read_tsv(my_file_function, show_col_types = FALSE)
#get all functions in separate cols and rows 
fun_genes <- pex_functions %>% 
  dplyr::select(symbol, geneFunction) %>% 
  separate_longer_delim(geneFunction, delim = "; ") 

#rowwise() %>% 
#  mutate(n_interactors = length(unlist(str_split(pex_interactor, ","))))

fun_genes_counted <- fun_genes %>% 
  count(geneFunction) %>% 
  arrange(desc(n))
```


Plot counts of gene function as bar plots
```{r}
fun_genes_counted %>% 
  ggplot(aes(y=reorder(geneFunction,n), x=n)) + 
  geom_bar(stat="identity", fill="#D3D3D3") +
  labs(y="") +
  theme_minimal()
```


Fig 1E: Comparison of network parameters (degree, betweenness centrality) of peroxins with other peroxisomal proteins
```{r}
#join degree tibble with function and plot boxgraph
#functions that should be compared
col_peroxins <- c("peroxin" = "#FDD49F", "other" = "#D9D9D9")

peroxins <- fun_genes %>% 
  filter(geneFunction == "peroxin") %>% 
  pull(symbol)

fun_pl <- fun_genes %>% 
  mutate(gene_class = if_else(symbol %in% peroxins, "peroxin", "other")) %>% 
  select(symbol, gene_class) %>% 
  unique() %>%
  left_join(pex_degree, by = join_by(symbol == symbol))

p <- fun_pl %>% 
  filter(!is.na(degree)) %>%
  ggplot(aes(x = gene_class, y=degree)) + 
  geom_boxplot(aes(fill = gene_class), stat="boxplot") + 
  scale_fill_manual(values = col_peroxins) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  labs(x = "") +
  stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 65, label = "p.format")

p
ggsave("output/figure1/degree_boxplot.pdf", plot = p, width = 3.5, height = 5.5, units = "cm", scale = 1.6)
```



```{r}
#plot betweenness centrality of peroxins vs other pex proteins
pex_between <- as_tibble(rownames_to_column(as.data.frame(betweenness(pex_graph, normalized = TRUE)), var="symbol")) %>% dplyr::rename("betweenness" = `betweenness(pex_graph, normalized = TRUE)`) 

fun_pl <- fun_genes %>% 
  mutate(gene_class = if_else(symbol %in% peroxins, "peroxin", "other")) %>% 
  select(symbol, gene_class) %>% 
  unique() %>%  
  left_join(pex_between, by = join_by(symbol == symbol))

p <- fun_pl %>% 
  filter(!is.na(betweenness)) %>%
  ggplot(aes(x = gene_class, y=betweenness)) + 
  geom_boxplot(aes(fill = gene_class), stat="boxplot") + 
  scale_fill_manual(values = col_peroxins) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  labs(x = "") +
  stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 0.4, label = "p.format")

p
ggsave("output/figure1/between_boxplot.pdf", plot = p, width = 3.5, height = 5.5, units = "cm", scale = 1.6)
```



```{r}
pex_closeness <- as_tibble(rownames_to_column(as.data.frame(closeness(pex_graph)), var="symbol")) %>% dplyr::rename("closeness" = `closeness(pex_graph)`)

fun_pl <- fun_genes %>% 
  mutate(gene_class = if_else(symbol %in% peroxins, "peroxin", "other")) %>% 
  select(symbol, gene_class) %>% 
  unique() %>%
  left_join(pex_closeness, by = join_by(symbol == symbol))

p <- fun_pl %>% 
  filter(!is.na(closeness)) %>%
  ggplot(aes(x = gene_class, y=closeness)) + 
  geom_boxplot(aes(fill = gene_class), stat="boxplot") + 
  scale_fill_manual(values = col_peroxins) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  labs(x = "") +
  stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 0.01, label = "p.format")

p
ggsave("output/figure1/closeness_boxplot.pdf", plot = p, width = 3.5, height = 5.5, units = "cm", scale = 1.6)


```


Fig 2: 
Homooliogmerisation
Compare homomers of UCRinteractome with homomers in ePI. Is there a statistical significant difference? 
```{r}
#UCRinteractome nodes 15631 (without ePI 15619)
#UCRinteractome selfloops 2137
#ePI nodes 173
#ePI selfloops (with iBRET) 78

xtab <- as.table(rbind(c(2137, 78), c(13494, 95)))
dimnames(xtab) <- list(homomers = c("Yes", "No"), interactome=c("UCR","ePI"))

library(rstatix)
homo_p <- fisher_test(xtab, detailed=TRUE)
homo_p
```

Comparison without bret data
```{r}
xtab_2 <- as.table(rbind(c(2137, 47), c(13494, 126)))
dimnames(xtab_2) <- list(homomers = c("Yes", "No"), interactome=c("UCR","ePI_withoutBRET"))
library(rstatix)

homo_p_withoutBRET <- fisher_test(xtab_2, detailed=TRUE)
homo_p_withoutBRET
```


Create Boxplot for Fig 2E: Homooligomerisation
```{r}
homo_plot <- tibble(freq = c(0.138, 0.450, 0.277), interactome = c("UCR", "ePI", "ePI_without_iBRET")) %>% 
  ggplot(aes(x = interactome, y = freq)) + 
  geom_bar(stat="identity", fill="#D3D3D3", color="black", width=0.7, linewidth=0.3) + 
  theme_minimal() + 
  theme(panel.border = element_rect(linetype = "blank", fill = NA),
        panel.grid = element_line(colour = "white"), 
        panel.grid.major.x = element_line(colour = "grey")) +
  labs(x = "", y = "") + 
  scale_y_continuous(limits = c(0,0.5)) +
  coord_flip()
homo_plot

ggsave("output/figure2/homomers_compare.pdf", plot = homo_plot, width = 5, height = 2.5, units = "cm", scale = 1.5)
```



Are non-peroxins enriched in homooligomers?
```{r}
ePI_edge_cl %>% filter(proteinA == proteinB) %>% mutate(function_class = if_else(proteinA %in% peroxins, "peroxin", "non-peroxin")) %>% group_by(function_class) %>% summarise(n=n())
```

```{r}
#peroxins n = 16
xtab_peroxin <- as.table(rbind(c(10, 65), c(6, 90)))
dimnames(xtab_peroxin) <- list(homomers = c("Yes", "No"), function_c=c("peroxin","non-peroxin"))

fisher_test(xtab_peroxin)
```


How is the ePI connected to the human interactome?

```{r}
#load data from Julia with internal / external degree per ePI protein
ePI_in_out_degree <- read_csv(my_intExt_degree_file) %>% 
  dplyr::rename("Entrez" = "...1", "degree" = "overall degree", "degree_int" = "internal degree", "degree_ext" = "external degree")
#merch with yifrach_genes
ePI_in_out_degree %>% left_join(yifrach_genes, by=join_by(symbol == Gene))
```
```{r}
require("ggrepel")
p<- ePI_in_out_degree %>% 
  left_join(yifrach_genes, by=join_by(symbol == Gene)) %>%
  filter(degree_int > 0) %>%
  ggplot(aes(x = degree_ext, y = FinalScore)) + 
  geom_point(alpha=0.5) + 
  geom_hline(yintercept = 8, linetype = "dashed") +
  geom_label_repel(aes(label = symbol),
                  size = 3,
                  box.padding = unit(0.35, "lines"),
                  point.padding = unit(0.3, "lines"), 
                  max.overlaps = 6) +
  labs(x = "external degree", y = "Yifrach score") + 
  theme_minimal() + 
  theme(panel.grid.major = element_line(colour = "grey"))

ggsave("output/figure2/degree_out.pdf", plot = p, width = 6, height = 3, units = "cm", scale = 2)
p

```


```{r}
require("ggrepel")
p<- ePI_in_out_degree %>% 
  left_join(yifrach_genes, by=join_by(symbol == Gene)) %>%
  filter(degree_int > 0) %>%
  ggplot(aes(x = degree_int, y = FinalScore)) + 
  geom_point(alpha=0.5) + 
  geom_hline(yintercept = 8, linetype = "dashed") +
  geom_label_repel(aes(label = symbol),
                  size = 3,
                  box.padding = unit(0.35, "lines"),
                  point.padding = unit(0.3, "lines"), 
                  max.overlaps = 6) +
  labs(x = "internal degree", y = "Yifrach score") + 
  theme_minimal() + 
  theme(panel.grid.major = element_line(colour = "grey"))

ggsave("output/figure2/degree_in.pdf", plot = p, width = 6, height = 3, units = "cm", scale = 2)
p
```

```{r}

require("ggrepel")
p<- ePI_in_out_degree %>% 
  left_join(yifrach_genes, by=join_by(symbol == Gene)) %>%
  filter(degree > 0) %>%
  ggplot(aes(x = degree, y = FinalScore)) + 
  geom_point(alpha=0.5) + 
  geom_hline(yintercept = 8, linetype = "dashed") +
  geom_label_repel(aes(label = symbol),
                  size = 3,
                  box.padding = unit(0.35, "lines"),
                  point.padding = unit(0.3, "lines"), 
                  max.overlaps = 6) +
  labs(x = "degree", y = "Yifrach score") + 
  theme_minimal() + 
  theme(panel.grid.major = element_line(colour = "grey"))

ggsave("output/figure2/degree_yifrach.pdf", plot = p, width = 6, height = 3, units = "cm", scale = 2)
p
```

