---
title: 'Figure 3: examine tissue specific networks'
output:
  html_document:
    df_print: paged
---


```{r}
library(tidyverse)
library(readxl)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(ggpubr)
```

Set paths to files
```{r}
dir <- ""

#load data for tissue expression per protein
protein_expr <- read_xlsx("data/HPA_PEX_ExpressedTiP_Matrix_transpose.xlsx") %>% 
  rename("...1" = "protein") %>%
  pivot_longer(cols = 2:51, names_to = "tissue", values_to = "expression")

tip_tissue <- read_xlsx("data/number_of_TiP_per_tissue.xlsx", col_names=c("tissue","n"), skip=1)


###### import information on peroxins ######
my_file_function <- "data/9_genesFunction.txt"  #functions of peroxisomal proteins
#read genes with their respective functions and sort
pex_functions <- read_tsv(my_file_function, show_col_types = FALSE)

#get all functions in separate cols and rows 
fun_genes <- pex_functions %>% 
  dplyr::select(symbol, geneFunction) %>% 
  separate_longer_delim(geneFunction, delim = "; ") 

peroxins <- fun_genes %>% 
  filter(geneFunction == "peroxin") %>% 
  pull(symbol)

col_peroxins <- c(peroxin = "#FDD49F", other = "#D9D9D9")

```

How often are proteins not expressed in any of the 50 tissues
```{r}
protein_expr %>% 
  group_by(protein, expression) %>% 
  summarise(n_notexpressed = n()) %>% 
  mutate(perc = n_notexpressed/sum(n_notexpressed)) %>% 
  filter(is.na(expression)) %>% 
  ungroup() %>% 
  dplyr::select(protein,n_notexpressed, perc) %>% 
  arrange(desc(perc)) %>% 
  print(n=100)
```


Fig 3B: Fraction of pex proteins expressed in tissue (expressed + TiP)
```{r}
p <- protein_expr %>% 
  filter(!is.na(expression)) %>% 
  group_by(tissue, expression) %>% 
  summarise(n = n()) %>% 
  mutate(fraction = n/sum(n)) %>% 
  mutate(n = if_else(expression == "expressed", sum(n), n)) %>% 
  mutate(fraction = if_else(expression == "expressed", n/172, fraction)) %>%
  mutate(sorting = if_else(expression == "expressed", n, 0)) %>%
  arrange(tissue, desc(expression)) %>%
  ungroup() %>%
  complete(tissue, expression, fill=list(n=0, fraction=0, sorting=0)) %>%
  ggplot(aes(x=reorder(tissue, -sorting), y = fraction, fill=expression)) +
  geom_bar(stat = "identity", position = position_identity()) +
  #geom_text(
  #  aes(label = n, group = expression),nudge_x = 0.1,
  #  color = "white"
  #) + 
  scale_fill_manual(values = c("#9A9999", "#E5D542")) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_y_continuous(name = "", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_x_discrete(name="") + 
  theme(legend.position="none")

print(p)
ggsave("output/figure3/fraction_expression_per_tissue.pdf", plot = p, width = 7, height = 4, units = "cm", scale = 2)


```
Fig. 3C: Are peroxin less often TiP than other peroxisomal proteins?
```{r}
p<- protein_expr %>% 
  filter(expression == "TiP") %>% 
  group_by(protein) %>% 
  summarise(n=n()) %>% 
  left_join(fun_genes %>% filter(geneFunction == "peroxin"), by=join_by(protein == symbol)) %>% 
  dplyr::select(protein, n, geneFunction) %>% 
  replace_na(list(geneFunction = "other")) %>% 
  ggplot(aes(x=geneFunction, y=n)) + 
  geom_boxplot(aes(fill = geneFunction), stat="boxplot") + 
  scale_fill_manual(values = col_peroxins) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  scale_y_continuous(breaks=c(0,2,4,6,8,10)) +
  labs(x = "", y = "n tissues") +
  stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 9, label = "p.format")

print(p)

ggsave("output/figure3/n_tissues_TiP_peroxins.pdf", plot = p, width = 5, height = 4, units = "cm", scale = 2)


```
Revision Analysis of n TiP for lysosomes
```{r}
#data files
lys_func <- read_xlsx("data/revision/proteins_GO_lysosome_20326.xlsx",sheet="overview")
lys_tip <- read_xlsx("data/revision/proteins_expert_curated_lysosome_TiP.xlsx")

p_lys <- lys_tip %>% 
  rowwise() %>% 
  mutate(n = length(unlist(str_split(TiP_in_tissue, ",")))) %>% 
  ungroup %>% 
  left_join(lys_func, by=join_by(GeneID == GeneID)) %>% 
  filter(!is.na(Function)) %>%
  ggplot(aes(x=Function, y=n)) + 
  geom_boxplot(aes(fill = Function), stat="boxplot") + 
  scale_fill_manual(values = c("Biogenese" = "#FDD49F", "Enzyme" = "#D9D9D9")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  scale_y_continuous(breaks=c(0,2,4,6,8,10)) +
  labs(x = "", y = "n tissues") +
  stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 9, label = "p.format")

print(p_lys)
```


Fig 3D: TiP LCC
```{r}
tip_genes_per_tissue <- read_xlsx("data/LCC_TiP_PEX.xlsx") %>% 
  filter(TiP_LCC_genes != "all single") %>% 
  mutate(TiP_LCC_genes = str_remove_all(TiP_LCC_genes, "[{|}| ]")) %>% 
  separate_longer_delim(TiP_LCC_genes, ",")

tip_lcc_pertissue <- tip_genes_per_tissue %>% 
  left_join(AnnotationDbi::select(org.Hs.eg.db, keys = unique(tip_genes_per_tissue$TiP_LCC_genes), keytype = "ENTREZID", columns = c("SYMBOL","ENTREZID")), by = c("TiP_LCC_genes" = "ENTREZID"))
```

Which TiP are most often member of a lcc
```{r}
#tissues that have sig larger lcc than expected by chance
#tissues lcc > 2
sig_tissues <- c("tongue", "parathyroid gland", "liver", "adrenal gland", "kidney", "duodenum", "midbrain", "adipose tissue", "skeletal muscle")

tip_lcc_pertissue %>% 
  filter(tissue %in% sig_tissues) %>% 
  group_by(SYMBOL) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
```


Figure 3D: Extract function of members of TiP-lcc

Function that takes fun_genes and a gene_list and outputs n of function per gene of gene_list
```{r}
getFunction <- function(fun_genes, gene_list) {
  fun_counts <- fun_genes %>% filter(symbol %in% gene_list) %>% count(geneFunction)
  return(fun_counts)
}
```

Get functions by tissue.
```{r}
#colors
tissue_filt = "tongue"
fun_cols <- deframe(read_tsv("data/colors/fun_colors.txt", show_col_types = FALSE))
#get barplots
# col fill "#D3D3D3"
gene_list <- tip_lcc_pertissue %>% filter(tissue == tissue_filt) %>% pull(SYMBOL)
p <- getFunction(fun_genes, gene_list) %>% 
  ggplot(aes(x = n, y = reorder(geneFunction, n), fill=geneFunction)) + 
  geom_bar(stat = "identity", color="black", width=0.9, linewidth=0.1) +
  theme_minimal() + 
  theme(panel.border = element_rect(linetype = "blank", fill = NA),
        panel.grid = element_line(colour = "white"), 
        panel.grid.major.x = element_line(colour = "grey")) +
  scale_x_continuous(breaks = seq(0,20,1)) +
  scale_fill_manual(values = fun_cols) +
  labs(x = "", y = "") +
  guides(fill = "none")

#ggsave
file_save = paste("output/figure3/",tissue_filt,"_function.pdf", sep="")
ggsave(file_save, plot = p, width = 4, height = 3, units = "cm", scale = 2)
```
