---
title: "expressionEnrichment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}
#load libraries
library(tidyverse)
library(readxl)
library(GO.db)
library(org.Hs.eg.db)
library(rstatix)
library(igraph)
```

Set files
```{r}
dir <- "../Manuskript/"

#tipNeighbors from Python Notebook
my_tipNeighbors_file <- paste(dir, "Julia_Analysen/173_ePI/results/parameter/TiPneighbors_to_PEX_LCC_alltissues_customsymbol.xlsx", sep="")
tipNeighbors <- read_xlsx(my_tipNeighbors_file)

```

Settings for GO enrichment and heatmap generation
```{r}
#GO enrichment
#set the desired GO-ontology here eg MF, BP or CC
ontology <- "CC"
pattern <- "(TiPneighbors_Julia)"
my_folder <- paste(dir, "Julia_Analysen/173_ePI/results/parameter/enrichment/Julia/", ontology, sep = "")
my_heatmap_file <- paste(dir, "Figures/Figure4/expressionEnrichmentHeatmap_",ontology,".pdf", sep = "")

```


Defines functions that read each enrichment file in a specific folder, concats all files in one big table and adds one extra column that specifies the tissue name. 
```{r }
read_concat_files = function(myFolder, pattern) {
  filenames <- list.files(myFolder, pattern=pattern, full.names=TRUE)
  tibs <- lapply(filenames, read_xls_files)
  for(i in 1:(length(tibs)-1)) { 
    tibs[[i+1]] = dplyr::bind_rows(tibs[[i]], tibs[[i+1]])
    }
  return(tibs[[length(tibs)]])
}

#provides functionality to read multiple tissue enrichment files
read_xls_files = function(file) {
  f <- str_split(file, "/", simplify = TRUE)
  n <- length(f)
  df <- read_excel(file, col_types= c("text","text","numeric","numeric","numeric","numeric","numeric","numeric","numeric")) %>% mutate(tissue = str_split(f[n], "_TiP", simplify = TRUE)[1])
  return(df)
}

```


Get the matrix with all TiP neighbors per tissue (exported from the python notebook)
TiPneighbors_to_PEX_alltissues_custom.csv

```{r}
#turn data into tidy format and remove all tissues with NA
tipNeighbors <- rename(tipNeighbors, "...1" = "TiP") %>% 
  pivot_longer(cols = 2:51, names_to = "tissue", values_to = "pex_interactor") %>% 
  filter(!is.na(pex_interactor)) %>% 
  mutate(pex_interactor = str_replace_all(pex_interactor,"\\['|'\\]|'","")) 

#a list of TiP and count how often they are TiP and neighbor of pex proteins
tipNeighbors %>% 
  count(TiP) %>% 
  arrange(desc(n)) %>%
  filter(n > 9)

print(tipNeighbors)
```

Save as supplementary file
```{r}
write_tsv(tipNeighbors, "10_nonPEX_TiPneighbors.txt")
```


Plot n of TiP neighbors per tissue
```{r}
bp <- tipNeighbors %>% 
  count(tissue) %>% 
  ggplot(aes(x = reorder(tissue, dplyr::desc(n)), y = n)) + 
  geom_bar(stat = "identity") + 
  theme_light() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),panel.grid.major.x = element_blank(), panel.grid.major.y = element_line(colour = "#F5F5F5", linetype ="dotted")) +
  xlab("") + 
  ylab("n of interacting TiP")

my_file <- paste(dir, "Figures/TiPperTissue.pdf", sep="")
ggsave(my_file, plot = bp)

bp

```

```{r}
#show n of TiPs 
tipNeighbors %>% 
  rowwise() %>% 
  mutate(n_interactors = length(unlist(str_split(pex_interactor, ",")))) %>% 
  ungroup %>% 
  arrange(desc(n_interactors))

print(tipNeighbors)
```


Plot enriched GO Terms of neighboring TiPs

Run functions to prepare data for plotting
```{r warning= FALSE}
dat_enr <- read_concat_files(my_folder, pattern)

```

Run a clustering to reorder tissues
```{r}
#filter for sign results and show only those terms that have > 1 genes 
dat_enr %>% 
  filter(`P-value-corrected`< 0.05) %>% 
  filter(Abundance_group > 1) %>% 
  dplyr::select(tissue, Association_term, `P-value-corrected`, OR) -> clustdata

clustdata %>% 
  dplyr::select(tissue, Association_term, `P-value-corrected`) -> temp
#prepare tibble for conversion into matrix
temp %>% 
  pivot_wider(names_from=c(Association_term), values_from= c('P-value-corrected')) -> clustdata_wide
rm(temp)
clustdata_wide %>% 
  mutate_all(~replace(.,is.na(.), 0)) -> temp
#convert to matrix
as.matrix(column_to_rownames(temp, var = "tissue")) -> matrix
rm(temp)

#cluster tissues
ord <- hclust( dist(matrix, method = "euclidean"))$order
#ord
#cluster go terms
ord_go <- hclust( dist(t(matrix), method = "euclidean"))$order
```



Now we plot the heatmap and save it to file "my_heatmap_file" as pdf
```{r}
# defines colours used in the heatmap -> color brewer but reversed order, since we want lower p-values to be "blueer"
cols = rev(RColorBrewer::brewer.pal(9, "Reds"))
#remove white from the cols and add a light-grey
cols = cols[-9]
cols[9] = "#f0f0f0" 

p <- clustdata %>% 
  mutate(Association_term = factor(Association_term, levels = colnames(matrix)[ord_go])) %>% 
  mutate(tissue = factor(tissue, levels = rownames(matrix)[ord])) %>%
  mutate(OR = if_else(is.infinite(OR), max(OR[is.finite(OR)]), OR)) %>%
  ggplot(aes(x=tissue, y=Association_term)) + 
  geom_tile(fill="white") + 
  theme_light() + 
  theme(panel.spacing = unit(0.25, "lines"),axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),panel.grid.major.y = element_line(colour = "grey", linetype ="dotted")) + 
  geom_point(aes(size = OR, fill = `P-value-corrected`, colour = `P-value-corrected`), alpha = 0.7) +
  scale_colour_gradientn(colors = cols, limits = c(0,0.05)) +
  scale_size(range = c(1,6)) +
  coord_flip() +
  labs(x = "", y = "") + 
  theme(legend.position="none")

#ggsave(my_heatmap_file, plot = p, width = 16, height = 12, unit="cm", scale = 1.8)

p
```

Write enrichment to file
```{r}
file_enr_to <- paste("GO_enrichment_TiP_neighbors_", ontology, ".txt", sep="")
write_tsv(clustdata, file = "file_enr_to")
```



Function that retrieves ppi between pex LCC members and non-pex TiP that are associated to a given GOID
```{r}

retrieveGenesForGOTerm <- function(term) {
  #get GOID
  goid <- AnnotationDbi::select(GO.db, keys = term, keytype = "TERM", columns="GOID")$GOID

  #retrieve gene names for GOID
  go_genes <- AnnotationDbi::select(org.Hs.eg.db, keys = goid, keytype = "GOALL", columns= "SYMBOL") %>% distinct(SYMBOL) %>% pull(SYMBOL)
  
  return(go_genes)
}

```


Set the term you want to search for and the tissue in which you want to search here: 
```{r}
term = "ciliary basal body"
tissue_to_search = "retina"

g <- retrieveGenesForGOTerm(term)
#show pex interactors (from tibble tipNeighbors)

neighbors_enr <- tipNeighbors %>% filter(TiP %in% g) %>% filter(tissue == tissue_to_search) %>% pull(TiP)
```

Get tissue-specific subnetwork of enriched TiP neighbors
```{r}

#set the file for the tissue specific interactome (in our case retina)
tissueFile <- "../Manuskript/Julia_Analysen/173_ePI/results/interactome_edgelist_retina.csv"
#genes for which we want to create a subnetwork
genes_sub <- c(neighbors_enr, c("CRAT", "PEX19", "RHOA", "VAPA", "MYO5A", "ACTG1", "AKAP11", "TRIM37", "IMPDH2", "PIPOX", "CCDC33", "VIM", "PEX13", "EEHADH")) 

tissue_neighbors_file <- paste(dir, "Figures/Figure4/retina_TiP_enr_neigbors.txt", sep="")
write_tsv(as.data.frame(neighbors_enr), tissue_neighbors_file)

interactome <- read_csv(tissueFile, skip=1, col_names=c("p1_entrez","p2_entrez","protein1","protein2"), col_types= "cccc") %>% 
  dplyr::select(p1_entrez,p2_entrez)

#get all unique entrezids
entrez_ids <- as.character(unique(c(unique(interactome$p1_entrez), unique(interactome$p2_entrez))))
#mapp names to entrezids
inter<- interactome %>% 
  left_join(AnnotationDbi::select(org.Hs.eg.db, keys = entrez_ids, keytype = "ENTREZID", columns = c("SYMBOL")), by = join_by(p1_entrez == ENTREZID)) %>% 
  left_join(AnnotationDbi::select(org.Hs.eg.db, keys = entrez_ids, keytype = "ENTREZID", columns = c("SYMBOL")), by = join_by(p2_entrez == ENTREZID)) %>% 
  dplyr::select(SYMBOL.x, SYMBOL.y) %>% 
  filter(!is.na(SYMBOL.x)) %>%
  filter(!is.na(SYMBOL.y))

network <- simplify(graph_from_data_frame(inter, directed = FALSE))
v_network <- intersect(V(network)$name, genes_sub)
tissue_network <- induced_subgraph(network, v_network)

#write to subnetwork file
subnetwork_file <- paste(dir, "Figures/Figure4/retina_enriched_network.txt", sep ="")
as.data.frame(as_edgelist(tissue_network)) %>% write_tsv(subnetwork_file)
```

