---
title: "Fig 1+2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read file with all positives (Cheng + BRET) and prepare data
```{r}
library(tidyverse)
library(readxl)
library(ggpubr)
library(rstatix)

#colors from iwanthue https://medialab.github.io/iwanthue/?utm_source=Nature+Briefing&utm_campaign=e0c3e024d4-briefing-dy-20211006&utm_medium=email&utm_term=0_c9dfd39373-e0c3e024d4-44542065
colors <- read_tsv("colors_fancy.txt", col_names = F, show_col_types = FALSE) %>% pull()
mygrey <- "#F5F5F5"
#output
my_heatmap_file = "../Manuskript/Figures/Figure2/2B_heatmap_positives.pdf"
#import
my_file <- "../Manuskript/Julia_Analysen/final_2023_10/results/Peroxisomal_interactions_across_dbs_final.xlsx" #file that contains comparison of lit vs bret interactions
my_file_mapping <- "../Manuskript/Julia_Analysen/results/translations/mapper.xlsx"
my_file_function <- "../Manuskript/Tables/genesFunction.txt"
```

Read functions 
```{r}
#read genes with their respective functions and sort
functions <- read_tsv(my_file_function, show_col_types = FALSE)

functions %>% arrange(desc(genefunction), symbol) %>% pull(symbol) -> genes_sorted_by_function
```


Count genes from the original results table 20211227_positivesBRET.xlsx
```{r}
positives_original <- read_tsv("../Manuskript/Julia_Analysen/positiveScreenAll_20181005_corrected.txt", col_names=FALSE, show_col_types = FALSE)
#following genes removed from analyses since there is no evidence for peroxisomal location (see also file "Nachfrage_pexProteine.xls")
remove_genes <- c("DNAJC10", "MPV17L2", "ACOT2", "ACOT2v1", "ACOT2v2", "MAVS", "RHOC", "SLC22A5", "ACSF3")
g <- union(unique(positives_original$X1), unique(positives_original$X2))
g_isoforms <- sort(setdiff(g, remove_genes))
g_isoforms
```

Prepare everything for heatmap plotting
```{r}
#read all positives
positives <- read_xlsx(my_file) %>% replace(is.na(.), 0)

#load mapper and prepare 
mapper <- read_excel(my_file_mapping)
mapper_red <- mapper %>% dplyr::select('Symbol[table]', 'EntrezID[mapper]') %>% group_by(`Symbol[table]`,`EntrezID[mapper]`) %>% distinct() %>% ungroup()

#map data
#code data
#found in ucrinteractome = 1
#found in iBRET = 2
#found in both = 3
#found nothing = 0
positives_mapped <- positives %>% left_join(mapper_red, by = c("Interactor_A" = "EntrezID[mapper]")) %>% left_join(mapper_red, by = c("Interactor_B" = "EntrezID[mapper]")) %>% arrange(Interactor_A) %>% dplyr::rename(proteinA = 'Symbol[table].x', proteinB = 'Symbol[table].y') %>% mutate(class = if_else(InteractomeUCR == 1,if_else(iBret == 1,3,1),if_else(iBret == 1,2,0)))

positives_mapped %>% group_by(class) %>% count(n())
```


```{r}
#do we have all proteins? n = 92 (without isoforms)
pg_withoutIso <- sort(union(unique(positives_mapped$proteinB), unique(positives_mapped$proteinA)))

pg_withoutIso
#clean up workspace
#rm(list = c("positives", "mapper", "functions", "my_file", "my_file_mapping", "my_file_function"))
```
lets get the isoforms included in the screen
```{r}
setdiff(g_isoforms, pg_withoutIso)
```

compare with yifrach dataset
```{r}
my_yifrach_genes <- "../Manuskript/Julia_Analysen/yifrach_bearbeitet.xlsx"
yifrach_genes <- read_excel(my_yifrach_genes)

yifrach_genes %>% mutate(study = if_else(Gene %in% pg_withoutIso,"y","n")) %>% group_by(study, DetectionFrequency) %>% count()
```

Fig 2b: Heatmap of ePI (iBRET positives plus literature interactions of all pex associated genes according to Yifrach et al)

(1) Read ePI interactions
```{r}
epi_file <- read_csv("../Manuskript/Julia_Analysen/final_2023_10/results/G_pex_edgelist.csv")
```


some operations that prepare the data for Fig 2B
specifically, this sets a lower triangle for heatmap plotting
```{r}
positives_mapped %>% dplyr::select(proteinA, proteinB, class) -> posA
positives_mapped %>% dplyr::select(proteinB, proteinA, class) %>% rename(proteinB = "proteinA", proteinA = "proteinB") -> posB

#bind everything together and create temp. dataframe. 
df = rbind(posA, posB)

#create a matrix out of the data frame
df_rearranged <- df %>% group_by(proteinA, proteinB) %>% unique %>% ungroup %>% arrange(proteinA) %>% pivot_wider(names_from = proteinA, values_from = class) %>% arrange(proteinB)

df_matr <- as.matrix(dplyr::select(df_rearranged,-proteinB)) 
rownames(df_matr) <- df_rearranged$proteinB

#convert back to data frame but only the lower triangle
ut <- lower.tri(df_matr, diag = TRUE)
df_lower <- data.frame(proteinB = rownames(df_matr)[row(df_matr)[ut]], proteinA = colnames(df_matr)[col(df_matr)[ut]], class = df_matr[ut])

#clean
rm(df)
```

Heatmap plotting
```{r}
p <- df_lower %>% 
  #filter(!is.na(class)) %>%
  mutate(class = case_match(class, 1 ~ "Literature", 2 ~ "iBRET", 3 ~ "Both", NA ~ "not interacting")) %>% 
  mutate(proteinBrev = factor(proteinB, levels = rev(unique(df_lower$proteinB)))) %>%
  ggplot() +
  geom_tile(aes(x = proteinA, y = proteinBrev, fill= class), color="white") +
  theme_light(base_size = 8) +
  scale_fill_manual(values=c("Both" = colors[5], "iBRET" = colors[4], "Literature" = colors[3], "not interacting" = mygrey), name="") +
    theme(
        panel.border = element_rect(linetype = "blank", fill = NA),
        panel.grid = element_line(colour = "white", linetype = "dotted"),
        panel.spacing = unit(0.25, "lines"),
        axis.text.x = element_text(
            angle = 90,
            hjust = 1,
            vjust = 0.5
        ),
    ) +
    labs(x = "", y = "") +
    guides(fill = "none")

ggsave(my_heatmap_file, plot = p, width = 12.5, height=12.5, scale = 1.3, units = "cm")
p
```
Fig 1D
Barplot Top 10 Hubs
```{r}
#exclude literature interactions
library(igraph)
pex_graph <- graph_from_edgelist(as.matrix(positives_mapped %>% filter(class != 1) %>% dplyr::select(proteinA, proteinB)), directed = FALSE)
pex_degree <- as_tibble(rownames_to_column(as.data.frame(degree(pex_graph, loops = FALSE)), var="symbol")) %>% rename("degree" = `degree(pex_graph, loops = FALSE)`) 

p_bar <- pex_degree %>%
  mutate(ranking = dense_rank(desc(degree))) %>%
  filter(ranking < 8) %>%
  ggplot(aes(x = reorder(symbol, degree), y = degree)) + 
  geom_bar(stat = "identity", fill="#D3D3D3", color="black", width=0.7, linewidth=0.3) + 
  theme_minimal() + 
  theme(panel.border = element_rect(linetype = "blank", fill = NA),
        panel.grid = element_line(colour = "white"), 
        panel.grid.major.x = element_line(colour = "grey")) +
  scale_y_continuous(limits = c(0,60)) +
  labs(x = "", y = "") + 
  coord_flip()
p_bar

ggsave("../Manuskript/Figures/Figure1/top_hubs.pdf", plot = p_bar, width = 3.5, height = 3.5, units = "cm", scale = 1.5)
```
Fig 1C degree distribution

```{r}
# List of degrees
G.degrees <- degree(pex_graph)

# Let's count the frequencies of each degree
G.degree.histogram <- as.data.frame(table(G.degrees))

# Need to convert the first column to numbers, otherwise
# the log-log thing will not work (that's fair...)
G.degree.histogram[,1] <- as.numeric(G.degree.histogram[,1])

# Now, plot it!
p <- ggplot(G.degree.histogram, aes(x = log(G.degrees), y = log(Freq))) +
  geom_point(shape = 21, colour = "black", fill = "grey", stroke = 1) +
  scale_x_continuous("Number of interaction partners") +
  scale_y_continuous("Number of proteins", breaks = c(0,1,2,3)) +
  theme_classic()
ggsave("../Manuskript/Figures/Figure1/degree_distribution.pdf", plot = p, width = 4, height = 3, units = "cm", scale = 2)
p

```

Load edgelist of extendedPI and plot degree distribution
```{r}

ePI_file <- ("../Manuskript/Julia_Analysen/final_2023_10/results/G_pex_edgelist.csv")
ePI_edge <- read_csv(ePI_file, show_col_types = FALSE)
ePI_graph <- graph_from_edgelist(as.matrix(ePI_edge %>% dplyr::select(Symbol_A, Symbol_B)), directed = FALSE)
# List of degrees
G.degrees <- degree(ePI_graph)

# Let's count the frequencies of each degree
G.degree.histogram <- as.data.frame(table(G.degrees))

# Need to convert the first column to numbers, otherwise
# the log-log thing will not work (that's fair...)
G.degree.histogram[,1] <- as.numeric(G.degree.histogram[,1])

# Now, plot it!
p <- ggplot(G.degree.histogram, aes(x = log(G.degrees), y = log(Freq))) +
  geom_point(shape = 21, colour = "black", fill = "grey", stroke = 1) +
  scale_x_continuous("Number of interaction partners") +
  scale_y_continuous("Number of proteins", breaks = c(0,1,2,3)) +
  theme_classic()
#ggsave("../Manuskript/Figures/Figure2/degree_distributio_ePI.pdf", plot = p, width = 4, height = 3, units = "cm", scale = 2)
p
```





Plot betweeness vs degree
```{r}
require("ggrepel")

p <- rownames_to_column(as.data.frame(cbind(betweenness(pex_graph, normalized = TRUE), degree(pex_graph, loops=FALSE))), var="symbol") %>% 
  ggplot(aes(x = V1, y = V2)) + 
  geom_point(alpha=0.5) + 
  geom_label_repel(aes(label = symbol),
                  size = 3,
                  box.padding = unit(0.35, "lines"),
                  point.padding = unit(0.3, "lines"), 
                  max.overlaps = 40) +
  labs(x = "betweeness centrality", y = "degree") + 
  theme_minimal() + 
  theme(panel.grid.major = element_line(colour = "grey"))

p

```




Analyse ppi between different functions
```{r}
#get all functions in separate cols and rows 
fun_genes <- functions %>% 
  dplyr::select(symbol, genefunction) %>% 
  separate_longer_delim(genefunction, delim = "; ") 

fun_genes %>% 
  count(genefunction) %>% 
  arrange(desc(n)) %>% 
  print(n=30)
```

Fig 1E: Comparison of network parameters (degree, betweenness centrality) of peroxins with other peroxisomal proteins
```{r}
#join degree tibble with function and plot boxgraph
#functions that should be compared
col_peroxins <- c("peroxin" = "#FDD49F", "other" = "#D9D9D9")

peroxins <- fun_genes %>% filter(genefunction == "peroxin") %>% pull(symbol)

fun_pl <- fun_genes %>% 
  mutate(gene_class = if_else(symbol %in% peroxins, "peroxin", "other")) %>% 
  select(symbol, gene_class) %>% 
  unique() %>%
  left_join(pex_degree, by = join_by(symbol == symbol))

p <- fun_pl %>% 
  ggplot(aes(x = gene_class, y=degree)) + 
  geom_boxplot(aes(fill = gene_class), stat="boxplot") + 
  scale_fill_manual(values = col_peroxins) +
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  labs(x = "") +
  stat_compare_means(method = "kruskal.test", label.x = 1.5, label.y = 65, label = "p.format")

p
ggsave("../Manuskript/Figures/Figure1/degree_boxplot.pdf", plot = p, width = 3.5, height = 5.5, units = "cm", scale = 1.6)
```
```{r}
#plot betweenness centrality of peroxins vs other pex proteins
pex_between <- as_tibble(rownames_to_column(as.data.frame(betweenness(pex_graph, normalized = TRUE)), var="symbol")) %>% rename("betweenness" = `betweenness(pex_graph, normalized = TRUE)`) 

fun_pl <- fun_genes %>% 
  mutate(gene_class = if_else(symbol %in% peroxins, "peroxin", "other")) %>% 
  select(symbol, gene_class) %>% 
  unique() %>%  
  left_join(pex_between, by = join_by(symbol == symbol))

p <- fun_pl %>% 
  ggplot(aes(x = gene_class, y=betweenness)) + 
  geom_boxplot(aes(fill = gene_class), stat="boxplot") + 
  scale_fill_manual(values = col_peroxins) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  labs(x = "") +
  stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 0.4, label = "p.format")

p
ggsave("../Manuskript/Figures/Figure1/between_boxplot.pdf", plot = p, width = 3.5, height = 5.5, units = "cm", scale = 1.6)
```
```{r}
pex_closeness <- as_tibble(rownames_to_column(as.data.frame(closeness(pex_graph)), var="symbol")) %>% rename("closeness" = `closeness(pex_graph)`)

fun_pl <- fun_genes %>% 
  mutate(gene_class = if_else(symbol %in% peroxins, "peroxin", "other")) %>% 
  select(symbol, gene_class) %>% 
  unique() %>%
  left_join(pex_closeness, by = join_by(symbol == symbol))

p <- fun_pl %>% 
  ggplot(aes(x = gene_class, y=closeness)) + 
  geom_boxplot(aes(fill = gene_class), stat="boxplot") + 
  scale_fill_manual(values = col_peroxins) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  labs(x = "") +
  stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 0.01, label = "p.format")

p
ggsave("../Manuskript/Figures/Figure1/closeness_boxplot.pdf", plot = p, width = 3.5, height = 5.5, units = "cm", scale = 1.6)


```
```{r}


#analyse again but different groups

groups_tofilter <- c("peroxin", "alpha oxidation", "beta oxidation", "ROS metabolism", "proliferation", "bile acid synthesis", "cholesterol/isoprenoid biosynthesis")

fun_pl <- fun_genes %>% 
    mutate(gene_class = case_when(genefunction == "peroxin" ~ "peroxin",
                                  genefunction == "alpha oxidation" ~ "alpha oxidation",
                                  genefunction == "beta oxidation" ~ "beta oxidation",
                                  genefunction == "proilferation" ~ "proliferation",
                                  genefunction == "bile acid synthesis" ~ "bile acid synthesis")) %>% 
    left_join(pex_degree, by = join_by(symbol == symbol))

p <- fun_pl %>% 
  filter(!is.na(gene_class)) %>% 
  ggplot(aes(x = gene_class, y=degree)) + 
  geom_boxplot(aes(fill = gene_class), stat="boxplot") + 
  #scale_fill_manual(values = col_peroxins) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  labs(x = "") +
  stat_compare_means(method = "kruskal.test", label.x = 2.5, label.y = 60, label = "p.format")

p


```



```{r}
#filter homomers from positives and literature interactions
pos_filter <- positives_mapped %>% 
  filter(class != 1) %>% 
  filter(proteinA != proteinB) %>% 
  dplyr::select(proteinA, proteinB)

#join with gene functions
pos_fun_joined <- pos_filter %>% 
  inner_join(fun_genes, by = join_by(proteinA == symbol), relationship = "many-to-many") %>% 
  inner_join(fun_genes, by = join_by(proteinB == symbol), relationship = "many-to-many")

pos_fun_joined %>%
  unite(col = "ppi_function", genefunction.x:genefunction.y, sep = "-") %>%
  count(ppi_function) %>% 
  arrange(desc(n))

```




