---
title: "examine tissue specific networks"
output: html_notebook
---


```{r}
library(tidyverse)
library(readxl)
library(ggdendro)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(ggpubr)

```

Set paths to files
```{r}
dir <- "../Manuskript/"

#load data for tissue expression per protein
protein_expr <- read_xlsx("../Manuskript/Julia_Analysen/173_ePI/results/parameter/HPA_PEX_ExpressedTiP_Matrix_transpose.xlsx") %>% 
  rename("...1" = "protein") %>%
  pivot_longer(cols = 2:51, names_to = "tissue", values_to = "expression")

tip_tissue <- read_xlsx("../Manuskript/Julia_Analysen/173_ePI/results/number_of_TiP_per_tissue.xlsx", col_names=c("tissue","n"), skip=1)
```

How often are proteins not expressed in any of the 50 tissues
```{r}
protein_expr %>% 
  group_by(protein, expression) %>% 
  summarise(n_notexpressed = n()) %>% 
  mutate(perc = n_notexpressed/sum(n_notexpressed)) %>% 
  filter(is.na(expression)) %>% 
  ungroup() %>% 
  dplyr::select(protein,n_notexpressed, perc) %>% 
  arrange(desc(perc)) %>% 
  print(n=100)
```


Fig 3B: Fraction of pex proteins expressed in tissue (expressed + TiP)
```{r}
p <- protein_expr %>% 
  filter(!is.na(expression)) %>% 
  group_by(tissue, expression) %>% 
  summarise(n = n()) %>% 
  mutate(fraction = n/sum(n)) %>% 
  mutate(n = if_else(expression == "expressed", sum(n), n)) %>% 
  mutate(fraction = if_else(expression == "expressed", n/172, fraction)) %>%
  mutate(sorting = if_else(expression == "expressed", n, 0)) %>%
  arrange(tissue, desc(expression)) %>%
  ungroup() %>%
  complete(tissue, expression, fill=list(n=0, fraction=0, sorting=0)) %>%
  ggplot(aes(x=reorder(tissue, -sorting), y = fraction, fill=expression)) +
  geom_bar(stat = "identity", position = position_identity()) +
  #geom_text(
  #  aes(label = n, group = expression),nudge_x = 0.1,
  #  color = "white"
  #) + 
  scale_fill_manual(values = c("#9A9999", "#E5D542")) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_y_continuous(name = "", limits=c(0,1), breaks=c(0,0.2,0.4,0.6,0.8,1)) +
  scale_x_discrete(name="") + 
  theme(legend.position="none")

p
#ggsave("../Manuskript/Figures/Figure3/fraction_expression_per_tissue.pdf", plot = p, width = 7, height = 4, units = "cm", scale = 2)


```
Fig. 3C: Are peroxin less often TiP than other peroxisomal proteins?
```{r}
p<- protein_expr %>% 
  filter(expression == "TiP") %>% 
  group_by(protein) %>% 
  summarise(n=n()) %>% 
  left_join(fun_pl, by=join_by(protein == symbol)) %>% 
  dplyr::select(protein, n, gene_class) %>% 
  replace_na(list(gene_class = "other")) %>% 
  ggplot(aes(x=gene_class, y=n)) + 
  geom_boxplot(aes(fill = gene_class), stat="boxplot") + 
  scale_fill_manual(values = col_peroxins) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1,vjust = 1), legend.position = "none") + 
  scale_y_continuous(breaks=c(0,2,4,6,8,10)) +
  labs(x = "", y = "n tissues") +
  stat_compare_means(method = "wilcox.test", label.x = 1.5, label.y = 9, label = "p.format")

p

#ggsave("../Manuskript/Figures/Figure3/n_tissues_TiP_peroxins.pdf", plot = p, width = 5, height = 4, units = "cm", scale = 2)


```



Fig 3D: TiP LCC
```{r}
tip_genes_per_tissue <- read_xlsx("../Manuskript/Julia_Analysen/173_ePI/results/LCC_TiP_PEX.xlsx") %>% 
  filter(TiP_LCC_genes != "all single") %>% 
  mutate(TiP_LCC_genes = str_remove_all(TiP_LCC_genes, "[{|}| ]")) %>% 
  separate_longer_delim(TiP_LCC_genes, ",")

tip_lcc_pertissue <- tip_genes_per_tissue %>% 
  left_join(AnnotationDbi::select(org.Hs.eg.db, keys = unique(tip_genes_per_tissue$TiP_LCC_genes), keytype = "ENTREZID", columns = c("SYMBOL","ENTREZID")), by = c("TiP_LCC_genes" = "ENTREZID"))
```

Which TiP are most often member of a lcc
```{r}
#tissues that have sig larger lcc than expected by chance
#sig_tissues <- c("tongue", "parathyroid gland", "liver", "adrenal gland", "kidney", "duodenum", "midbrain", "choroid plexus", "spinal cord", "salivary gland", "heart muscle", "adipose tissue", "skeletal muscle", "colon",  "small intestine")
#tissues lcc > 2
sig_tissues <- c("tongue", "parathyroid gland", "liver", "adrenal gland", "kidney", "duodenum", "midbrain", "choroid plexus", "heart muscle", "adipose tissue", "skeletal muscle",  "small intestine")

tip_lcc_pertissue %>% filter(tissue %in% sig_tissues) %>% group_by(SYMBOL) %>% summarise(n = n()) %>% arrange(desc(n))
```

Matrix plot of TiP per Tissue

Cluster before plotting
```{r}

temp_clustering <- tip_lcc_pertissue %>% 
  filter(tissue %in% sig_tissues) %>%
  dplyr::select(tissue, SYMBOL) %>% 
  mutate(ppi = 1) %>% 
  complete(tissue, SYMBOL, fill=list(ppi = 0))

temp_matrix <- as.matrix(column_to_rownames(temp_clustering %>% pivot_wider(names_from = SYMBOL, values_from = ppi), var="tissue"))

clust_tissue <- hclust( dist(temp_matrix, method = "binary"))
clust_genes <- hclust( dist(t(temp_matrix), method = "binary"))
ord_tissues <- clust_tissue$order
dendro_tissue <- as.dendrogram(clust_tissue)
dendro_tissue_plot <- ggdendrogram(dendro_tissue, rotate=TRUE) + theme(axis.text.x = element_blank(), axis.text.y = element_blank())

ord_genes <- clust_genes$order
dendro_genes <- as.dendrogram(clust_genes)
dendro_genes_plot <- ggdendrogram(dendro_genes) + theme(axis.text.x = element_blank(), axis.text.y = element_blank())

tip_map <- temp_clustering %>% 
  mutate(SYMBOL = factor(SYMBOL, levels = colnames(temp_matrix)[ord_genes])) %>%
  mutate(tissue = factor(tissue, levels = rownames(temp_matrix)[ord_tissues])) %>%
  mutate(ppi = as.character(ppi)) %>% 
  ggplot() + 
  geom_tile(aes(x=tissue, y=SYMBOL, fill=ppi), color="white") +
  scale_fill_manual(values = c("#D9D9D9", "#E5D542")) + 
  theme_light() +
  theme(
        panel.border = element_rect(linetype = "blank", fill = NA),
        panel.grid = element_line(colour = "white", linetype = "dotted"),
        panel.spacing = unit(0.3, "lines"),
        axis.text.x = element_text(
            angle = 90,
            hjust = 1,
            vjust = 0.5),
        ) +
  scale_x_discrete(name="") + 
  scale_y_discrete(name="") +
  guides(fill = "none") +
  coord_flip()
tip_map
  
```


Save heatmap to files
```{r}
ggsave("../Manuskript/Figures/Figure3/tip_heatmap.pdf", plot = tip_map, width = 13, height = 5, units = "cm", scale = 1.5)
```



Extract tissue specific modules: function
```{r}
#load network and write submodule as csv
library(igraph)
extract_submodule <- function(tissueFile, genes) {
  interactome <- read_csv(tissueFile, skip=1, col_names=c("p1_entrez","p2_entrez","protein1","protein2"), show_col_types = FALSE) %>% dplyr::select(protein1,protein2)
  network <- simplify(graph_from_data_frame(interactome, directed = FALSE))
  v_network <- intersect(V(network)$name, genes)
  tissue_network <- induced_subgraph(network, v_network)

  return(tissue_network)
}
```

Extract TiP interactions
```{r}

folder_from <- "../Manuskript/Julia_Analysen/173_ePI/results/ePI_edgelist_"
folder_to <- "../Manuskript/Figures/Figure3/tissueNetworks/"
gene_sets <- list()

for (i in 1:length(sig_tissues)) {
  file <- paste(folder_from, sig_tissues[i], ".csv", sep="")
  genes <- tip_lcc_pertissue %>% filter(tissue == .env$sig_tissues[i]) %>% pull(SYMBOL)
  #add genes to our gene_sets list
  gene_sets[[sig_tissues[i]]] <- genes
  net <- extract_submodule(file, genes)
  #save edge list for further processing e.g. to construct network images
  file_to <- paste(folder_to, sig_tissues[i], sep="")
  as.data.frame(as_edgelist(net)) %>% write_tsv(file_to, col_names = FALSE)
}
```

Function of members of TiP-lcc
```{r}
my_file_function <- "../Manuskript/Tables/genesFunction.txt"
#read genes with their respective functions and sort
functions <- read_tsv(my_file_function, show_col_types = FALSE)
#get all functions in separate cols and rows 
fun_genes <- functions %>% 
  dplyr::select(symbol, geneFunction) %>% 
  separate_longer_delim(geneFunction, delim = "; ") %>%
  filter(!geneFunction %in% c("membrane", "matrix"))
```

Function that takes fun_genes and a gene_list and outputs n of function per gene of gene_list
```{r}
getFunction <- function(fun_genes, gene_list) {
  fun_counts <- fun_genes %>% filter(symbol %in% gene_list) %>% count(geneFunction)
  return(fun_counts)
}
```


```{r}
#colors
tissue_filt = "tongue"
fun_cols <- deframe(read_tsv("fun_colors.txt", show_col_types = FALSE))
#get barplots
# col fill "#D3D3D3"
gene_list <- tip_lcc_pertissue %>% filter(tissue == tissue_filt) %>% pull(SYMBOL)
p <- getFunction(fun_genes, gene_list) %>% 
  ggplot(aes(x = n, y = reorder(geneFunction, n), fill=geneFunction)) + 
  geom_bar(stat = "identity", color="black", width=0.9, linewidth=0.1) +
  theme_minimal() + 
  theme(panel.border = element_rect(linetype = "blank", fill = NA),
        panel.grid = element_line(colour = "white"), 
        panel.grid.major.x = element_line(colour = "grey")) +
  scale_x_continuous(breaks = seq(0,20,1)) +
  scale_fill_manual(values = fun_cols) +
  labs(x = "", y = "") +
  guides(fill = "none")

#ggsave
file_save = paste("../Manuskript/Figures/Figure3/",tissue_filt,"_function.pdf", sep="")
ggsave(file_save, plot = p, width = 4, height = 3, units = "cm", scale = 2)
```


Matrix degree of a protein by tissue
```{r}
#print matrix: pex proteins vs degree by different tissues
#read file
tissue_degree <- read_csv("../Manuskript/Julia_Analysen/173_ePI/results/parameter/_PEX_degree.csv", show_col_types = FALSE) 
tissue_degree_tidy <- tissue_degree %>%
  filter(all != 0) %>%
  dplyr::select(!c(all,GeneID)) %>%
  mutate_all(~replace(.,is.na(.), 0)) %>%
  pivot_longer(cols = 1:50, names_to = "tissue", values_to = "degree") %>% 
  group_by(Symbol) %>% 
  mutate(z_score = scale(degree)) %>% 
  ungroup %>% 
  mutate_all(~replace(.,is.na(.), 0)) #this is needed to circumvent NaN values in zscores

#perform clustering: 
#create data matrix
matrix_ctd <- as.matrix(column_to_rownames(tissue_degree_tidy %>% dplyr::select(!degree) %>% pivot_wider(names_from = tissue, values_from = z_score), var = "Symbol"))

#cluster genes
clust_genes <- hclust( dist(matrix_ctd, method = "euclidean"))
ord_genes <- clust_genes$order
dendro_genes <- as.dendrogram(clust_genes)
dendro_genes_plot <- ggdendrogram(dendro_genes) + theme(axis.text.x = element_blank(), axis.text.y = element_blank())

#cluster tissues
clust_tiss <- hclust( dist(t(matrix_ctd), method = "euclidean"))
ord_tiss <- clust_tiss$order
dendro_tiss <- as.dendrogram(clust_tiss)
dendro_tiss_plot <- ggdendrogram(dendro_tiss, rotate=TRUE) + theme(axis.text.x = element_blank(), axis.text.y = element_blank())

#defines colours used in the heatmap
cols = RColorBrewer::brewer.pal(9, "Blues")

p<- tissue_degree_tidy %>% 
  mutate(Symbol = factor(Symbol, levels = rownames(matrix_ctd)[ord_genes])) %>%
  mutate(tissue = factor(tissue, levels = colnames(matrix_ctd)[ord_tiss])) %>%
  ggplot(aes(Symbol, tissue)) + 
  geom_tile(aes(fill = z_score))+
  theme_light() + 
  theme(panel.spacing = unit(0.25, "lines"),axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),panel.grid.major.y = element_line(colour = "grey", linetype ="dotted")) + 
  scale_fill_gradientn(colors = (cols)) +
  labs(x = "", y = "")

# printing (cm)
A4width = 21
A4height = 29.7
  
ggsave("../Manuskript/Figures/Figure3/degree_by_tissue.pdf", plot = p, width = 0.4*A4width, height = 0.175*A4height, scale = 1.5)

p
```
